# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r,message=FALSE,, warning=FALSE}
require(data.table)
require(ggplot2)
stepdata <- fread("activity.csv")
stepdata$date <- as.Date(stepdata$date)
```


## What is mean total number of steps taken per day?
```{r, echo=FALSE, results='hide', warning=FALSE }

forplot <- as.data.table(stepdata[,lapply(.SD, function(x) sum(x, na.rm = TRUE)), by = "date", .SDcols = 1])

histog <- with(forplot,qplot(date,steps, stat = "identity", geom = "histogram"))

avg <- mean(forplot$steps)
med <- median(forplot$steps)

```

```{r, echo=FALSE}

print(histog)

```

The mean number of steps per day is `r avg` and the median number of steps per day is `r med`.




## What is the average daily activity pattern?
```{r, echo=FALSE, results='hide', warning=FALSE}
forplot2 <- as.data.table(stepdata[,lapply(.SD, function(x) mean(x, na.rm = TRUE)), by = "interval", .SDcols = 1])

line1 <- with(forplot2,qplot(interval, steps, geom = "line"))

maxinter <- forplot2$interval[which.max(forplot2$steps)]
```

```{r, echo=FALSE}

print(line1)

```


The 5-minute interval, on average across all days that contains the maximum number of steps is `r maxinter`.



## Imputing missing values

```{r, echo=FALSE, results='hide', warning=FALSE}

nas <- stepdata[!complete.cases(stepdata),.N]

```

The number of rows containg NAs, incomplete cases, is `r nas`.


To replace the missing data, I have decided to replace it with median rather than mean data.  Numbers based on the mean could be skewed by some abnormally high activity during any given 5-minute interval.  Median is not perfect, but is more likely to match the "typical" 5-minute interval.  Code below.

```{r, results='hide', warning=FALSE}
stepdata2 <- stepdata
replaceNAs <- as.data.table(stepdata2[,lapply(.SD, function(x) median(x, na.rm = TRUE)), by = "interval", .SDcols = 1])
stepdata2 <- merge(stepdata2, replaceNAs, by = "interval", all.x = TRUE)
sub <- subset(stepdata2, is.na(steps.x))
sub$steps.x <- sub$steps.y 
stepdata2 <- subset(stepdata2, !is.na(steps.x))
stepdata2 <- rbind(stepdata2, sub)
setnames(stepdata2, "steps.x", "steps")
stepdata2$steps.y <- NULL
```

```{r, echo=FALSE, results='hide', warning=FALSE }

forplot3 <- as.data.table(stepdata2[,lapply(.SD, function(x) sum(x, na.rm = TRUE)), by = "date", .SDcols = 2])

histog2 <- with(forplot3,qplot(date,steps, stat = "identity", geom = "histogram"))

avg2 <- mean(forplot3$steps)
med2 <- median(forplot3$steps)

avgchange <- avg2-avg
medchange <- med2-med


```

```{r, echo=FALSE}

print(histog2)

```

```{r, echo=FALSE, results='hide', warning=FALSE }

forplot4 <- merge(forplot, forplot3,by = "date")
forplot4$stepsdiff <- forplot4$steps.y -  forplot4$steps.x

histog3 <- with(forplot4,qplot(date,stepsdiff, stat = "identity", geom = "histogram"))

```

The new mean number of steps per day is `r avg` and the new median number of steps per day is `r med`.  The change to the mean is `r avgchange` steps and the change to the median is `r medchange` steps.  The impact of inputing missing is an increase to the total daily number of steps.  See below.

```{r, echo=FALSE}

print(histog3)

```

## Are there differences in activity patterns between weekdays and weekends?


```{r, echo=FALSE, results='hide', warning=FALSE }
stepdata2$daytype <- weekdays(stepdata2$date)
stepdata2$daytype[stepdata2$daytype == "Sunday"] <- "Weekend"
stepdata2$daytype[stepdata2$daytype == "Saturday"] <- "Weekend"
stepdata2$daytype[stepdata2$daytype != "Weekend"] <- "Weekday"
stepdata2$daytype <- as.factor(stepdata2$daytype)

line2 <- qplot(interval, steps,data = stepdata2, stat="summary", fun.y="mean", geom = "line") + facet_wrap(~daytype, nrow=2)

```

```{r, echo=FALSE}

print(line2)


